\chapter{Introduction}
\label{chapter:introduction}

% The page numbering must be reset here inside the file
\pagenumbering{arabic}
\setcounter{page}{1}


\section{Aim and scope of this document}
\label{sec:aim}
%
\smr\ performs passive measurements of the atmosphere, mainly at wavelengths of
$\approx$ 0.6\,mm. The basic observation of \smr\ is spectra in different
frequency bands. After calibration the spectra are denoted as L1b data, and are
grouped into limb scans for further processing. The overall aim of this
document is to describe the software and algorithms used for the L2 processing,
that is, the step of extracting geophysical data from the L1b. The aim includes
also to motivate the different decisions involved, such as what instrumental
parameters that shall be retrieved in parallel to the geophysical data to
provide best possible L2 data product. The document describes the retrieval
process and related questions in a general manner, while details and
information that varies between frequency bands are found elsewhere (see
Sec.~\ref{sec:reading}).

The document covers only the ``operational'' \smr\ L2 products. These products
are based on measurements confined to the meso- and stratosphere, down to a few
kilometres above the tropopause\todo{How and where will we set this limit?}, to
allow that scattering can be neglected in the simulations of atmospheric
radiative transfer. Observations that can be affected by scattering (due to
tropospheric clouds) are treated separately, and the same applies to data
measured when Odin has been performed special scanning sequences (see further
Sec.~\ref{sec:nonstandard}). The result of these retrievals are formally also
L2 products, but inside this document the term ``L2'' refers only to the
operational products.

A much more in-depth overview of the \smr\ L2 processing is found Chapter~2.
The software tool that handles the tasks directly associated with atmospheric
radiative transfer and sensor modelling is described in Chapter~3, while the
actual data extraction step is covered by Chapter~4. No L2 data are complete
without a careful characterisation of spatial resolution and errors, which is
the topic of Chapter~5. Finally, Chapter 6 gives a summary, with focus on the
most important points to correctly understand and apply the \smr\ L2 data.
A list of used acronyms is found directly after the table of contents.




\section{\smr}
\label{sec:odin}
%
\subsection{The Odin satellite}
%
The Odin satellite was launched on 20 February 2001, into a sun-synchronous
18:00 hour ascending node orbit, carrying two co-aligned limb sounding
instruments: OSIRIS (optical spectrograph and infrared imaging system) and
\SMR\ (sub-millimetre radiometer). Originally, Odin was used for both
atmospheric and astronomical observations, while since 2007 only its aeronomy
mission is active. Odin is a Swedish-led project, in cooperation with Canada,
France and Finland. Both of Odin's instruments are still functional, and the
present operation of the satellite is partly performed as a ESA third party
mission.


\subsection{The \SMR\ instrument}
%
The \smr\ package is highly tunable and flexible. In short, the four main
receiver chains can be tuned to cover frequencies in the ranges
486\,--\,504\,GHz and 541\,-- \,581\,GHz, but the maximum total instantaneous
bandwidth is only 1.6\,GHz. This bandwidth is determined by the two
auto-correlation spectrometers (ACs) used for atmospheric observations. The two
ACs can be coupled to any of the four front-ends, but only two or three
front-ends are used simultaneously. The ACs cover 400 or 800\,MHz per
front-end, depending on configuration. In the configuration applied for
atmospheric sounding, the channels of the ACs have a spacing of 1\,MHz, while
the frequency resolution is only 2\,MHz. To cover all molecular transitions of
interest, a number of ``observation modes'' has been defined. Each observation
mode makes use of two or three frequency bands. Single sideband operation is
obtained by tunable Martin-Pupplet interferometers. The nominal sideband
suppression is better than 19\,dB across the image band.

\smr\ has also a receiver chain around the 118\,GHz oxygen transition. This
receiver was heavily used during Odin's astronomy mission. For the atmospheric
mission, this front-end was planned to be used for retrieving temperature
profiles, but a technical problem (drifting LO frequency) and the fact that the
analyses requires treatment of Zeeman effects have given these data low
priority and are not yet processed properly.

The main reflector of \smr\ has a diameter of 1.1\,m, giving a
vertical resolution at the tangent point of about 2\,km. The vertical scanning
of the two instruments' line-of-sigh is achieved by a rotation of the satellite
platform, with a rate matching a vertical speed of the tangent altitude of
750\,m/s. Measurements are performed during both upward and downward scanning.
The lower end of the scan is typically at about 7\,km, the upper end varies
between 70\,km and 110\,km, depending on observation mode. In correspondence,
the horizontal sampling ranges from 1 scan per 600 km to 1 scan per 1000 km.
Measurements are in general performed along the orbit plane, providing a
latitude coverage between 82.5$^{\circ}$S and 82.5$^{\circ}$N. Since the end of
2004 Odin is also pointing off-track during certain periods, e.g.\ during the
austral summer season, allowing the latitudinal coverage to be extended towards
the poles. 


\subsection{Main instrumental error sources}
%
The receiver noise temperature of \smr\ differs between the front-ends, varying
between ? and ?\,K\todo{Add values}. This makes thermal noise an important
error source. The pointing of Odin can be reconstructed to an accuracy matching
?\,m in tangent altitude, and observed frequencies can be controlled with an
accuracy better than 2\,MHz. The retrieval process includes further estimation
of pointing and frequency, but there will be some remaining errors.

The three error sources discussed above basically follow nominal performance
and are well characterised, but there are also features that are not yet
understood. First of all, it is clear that the sideband suppression does not
reach the nominal value of 19\,dB in all frequency bands. Ongoing tests and
investigations aim to set better constrains on the actual suppression achieved.
Further, the sub-bands of the ACs are not always overlapping perfectly, and
some degree of non-linear response can not be ruled out. Both these issues are
most problematic when measuring spectral features giving a high range of
brightness temperatures over the frequency band.

There are also non-nominal features that are handled, or characterised, fairly
well. Reflections inside the receiver give rise to ``baseline ripple'', but
these ripples are largely removed as part of the L1b processing. The
calibration uncertainty increases linearly with measured brightness
temperature, and is presently estimated to not exceed 2\,K at 220\,K. \smr\
applies Dicke switching with respect to ``cold sky'' but there is some
remaining impact of gain variations. These variations give rise to a constant
shift of the brightness temperatures across the band (that is, it can be seen
as a flat baseline ripple), where the shift has a standard deviation of about
2\,K and is uncorrelated between tangent altitudes (and front-ends). The
retrieval can remove these shifts with marginal impact on the retrievals as
long as some part of the measured spectrum corresponds to cosmic background
radiation (that is, high tangent altitudes), but this is a main error source
for other situations.



\section{Further reading}
\label{sec:reading}
%
\citep{murtagh:anove:02} give an overview of the Odin aeronomy mission, as well
as some technical details of \smr. Further technical information is found in
e.g.\ \citet{eriksson:studi:02}, and a more detailed description of the
observation modes is provided by \citet{merino:studi:02}.

The input to this L2 processing is the L1b data described in ?\addref. The
exact settings applied and the general characteristics of the L2 data, that
both vary between the frequency bands, are described in ?\addref.





%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "L2_ATBD"
%%% End: 
