\chapter{Retrieval variables}
\label{chapter:retvars}

This chapter reviews the variables that are part of the retrievals, either
directly or indirectly. The presentation follows the formalism presented in
Sec.~\ref{sec:formalism}.


\section{Basic variable choices}
\label{sec:retr:basics}
%
\smr\ performs measurements in several frequency bands simultaneously
(Sec.~\ref{sec:smr:details}). The set of frequency bands differs between
observation modes\addref. The altitude range scanned by \smr\ is not fixed, and
a \smr\ ``scan'' is simply defined as the spectra recorded between two adjacent
turning points, in altitude, of the limb scanning\addref. The retrievals covered
by this \ATBD\ operate only on one scan and spectra from a single frequency
band at the time. Each combination of scan and frequency band is treated
separately, no information is carried between different retrievals.

Following the discussion in Sec.~\ref{sec:setup:inverse:xb}, the state vector
does not only contain the quantities targeted directly by the \smr\
measurements, that is, gas species concentrations and atmospheric temperatures,
but also a number of other variables are part of \SttVct\ (Sec.~\ref{sec:x}).
These variables are either instrument correction factors or describe some
interfering atmospheric effect. However, a restriction to effects of random
nature is made, no fixed forward model variables (such as, spectroscopic data)
are placed in \SttVct. That is, variables causing systematic retrieval errors
are not included in \SttVct\ (and neither in \aCvrMtr{o}, Sec.~\ref{sec:So}).

Covariance matrices are implemented as sparse matrices, while all other
variables are stored as full vectors or matrices. For the retrieval problem at
hand, $\aWfnMtrTrp{\SttVct}$ is relatively dense and the overhead related with
a sparse representation results in better performance keeping
$\aWfnMtrTrp{\SttVct}$ as a full matrix



\section{The measurement vector, \MsrVct}
\label{sec:y}
%
The measurement vector contains solely \smr\ brightness temperature values,
with spectra from different tangent altitudes appended. That is,
\begin{equation}
  \label{eq:y}
  \MsrVct =
  \begin{bmatrix}
    \MsrVct_1\\ 
    \MsrVct_2\\ 
    \vdots\\ 
    \MsrVct_s
  \end{bmatrix},
\end{equation}
where $\MsrVct_1$/$\MsrVct_x$ is the spectrum from the highest/lowest tangent
altitude etc. 

Each retrieval does only contain data from a single scan, as mentioned, but all
spectra of the scan are not included. For each mode an upper and lower
tangent altitude limit are defined, that determine which spectra of a scan that
will be part of \MsrVct. The altitude limits are defined using geometrical
tangent altitudes (that is, refraction neglected) as those are the tangent
altitudes found in the L1b data.

The tangent altitude limits are allowed to vary as a function of latitude, an
example is given in Fig.~?\todo{Make figure}. These tangent altitude masks are
throughout defined at 90\degree S, \dots\todo{Add values}\ and 90\degree N,
with linear interpolation applied between these latitudes. The altitude limits
applied for each frequency band are found ?\todo{Where?}.



\section{Observation uncertainty covariance matrix, \aCvrMtr{o}}
\label{sec:So}
%
All variables of random nature that have a significant impact on the
measurements, including indirect effects through the forward model, are part of
the state vector (Secs.~\ref{sec:retr:basics} and \ref{sec:x}). This means that
the term $\aWfnMtr{\FrwMdlVct}\aCvrMtr{\FrwMdlVct} \aWfnMtr{\FrwMdlVct}^T$ in
Eq.~\ref{eq:So} should give a relatively small contribution (ignoring
systematic components) to $\aCvrMtr{\MsrErrVct_o}$. Accordingly, \aCvrMtr{o} is
set to only represent thermal noise:
\begin{equation}
  \aCvrMtr{o} = \aCvrMtr{\MsrErrVct_n}.
\end{equation}
\\
\lcomment{PE}{Exactly what expression to use for converting L1b data to thermal
  noise standard deviation? Do we consider the noise correlation between
  channels? With Hanning the correlation is significant, while without Hanning
  a pure diagonal matrix is a better approximation.}


\section{The state vector, \SttVct}
\label{sec:x}
%
The state vector contains a mix of quantities (Sec.~\ref{sec:retr:basics}).
Besides the 118\,GHz band, the quantities always present are: gas species
profiles $(\VctStl{v})$, the temperature profile $(\VctStl{t})$, a set of
brightness temperature off-sets $(\Delta\VctStl{o})$, a pointing off-set
$(\Delta\theta)$ and a frequency off-set $(\Delta\nu)$\todo{Correct?}. With
just these quantities retrieved, and a case involving two gas species, the
state vector becomes:
\begin{equation}
  \label{eq:x:example}
  \SttVct =
  \begin{bmatrix}
    \VctStl{v}_1\\ 
    \VctStl{v}_2\\ 
    \VctStl{t}\\ 
    \VctStl{\Delta\VctStl{o}}\\ 
    \Delta\theta\\ 
    \Delta\nu 
  \end{bmatrix}.
\end{equation}
For 118\,GHz the state vector contains no species profiles as only oxygen,
having a known VMR, has to be considered for this frequency band.

A species or temperature ``profile'' is a set of values describing the vertical
variation, at a set of specified pressures, that is, the retrieval grid (see
also Sec.~\ref{sec:arts:aspects}). Such retrieval grids are discussed further
in Sec.~\ref{sec:b:pgrid}, together with the basic pressure grid of the forward
model. Temperatures are given in Kelvin.
\\
\lcomment{PE}{Do we go for log-retrievals for species? Discuss basic aspects of
  the log conversion here.}

As discussed in Sec.~\ref{sec:smr:errors}, there is not a perfect compensation
for gain variations, causing an off-set in brightness temperatures. This
off-set has been found to be flat over the frequency bands, but differ from one
tangent altitude to next. The vector \VctStl{\Delta\VctStl{o}} contains one
brightness temperature off-set for each included spectrum, to compensate for
the ``baseline shift''. That is, a correction of the brightness temperatures is
estimated, and corrected spectra can be expressed as
\begin{equation}
  \label{eq:retr:baseline}
  \hat{T}_b^{ij} = T_b^{ij} + \Delta\VctStl{o}_i
\end{equation}
where $T_b^{ij}$ is brightness temperature of channel $j$ and tangent altitude
$i$ and $\Delta\VctStl{o}_i$ is element $i$ of $\Delta\VctStl{o}$.

The pointing and frequency off-sets are scalar values. This implies that these
correction factors are valid for the complete scan. In \ARTS\ the viewing
direction, of each \LOS, is specified as the angle from zenith. These angles
can not be determined perfectly from attitude data, but the error has been
judged to be constant during a scan and a single correction value should be
sufficient. Hence, the retrieved zenith angles are
\begin{equation}
  \label{eq:retr:point}
  \hat{\theta_i} = \theta_i + \Delta\theta,  
\end{equation}
where $\theta_i$ is the L1b zenith angle of tangent altitude $i$. A positive 
$\Delta\theta$ means that the retrieval estimated lower tangent altitudes than
obtained by the attitude data. As a rule of thumb, $\Delta\theta=\pm0.01\degree$
matches about $\mp550$\,m in tangent altitude.

There is a similar uncertainty in the exact frequencies observed. This
uncertainty originates in the \LO-signal. Already pre-launch tests hows that
\LO-frequency varies somewhat with temperature. A correction table was created,
but some frequency error remains and there seems also to be some ageing on the
\LO-chain affecting the final frequencies. The final frequency uncertainty is
totally correlated between the backend channels inside each frequency band. As
long as the \LO-signal is phase-locked, it is judged that the frequency off-set
is constant during a limb scan. However, the phase-locking of the 118 and
575\,GHz front-ends have been failing basically throughout the mission, and the
assumption of a constant frequency off-set is more uncertain for these receiver
chains. The switch from version 1 to 2 of \ARTS\ enables the retrieval of a
non-constant frequency off-set, but this possibility will just be activated later
if found necessary. The retrieved frequency of channel $i$ is
\begin{equation}
  \label{eq:retr:freq}
  \hat{\nu_j} = \nu_j + \Delta\nu,  
\end{equation}
where $\nu_j$ is the frequency found given by L1b data.\\
\\
\lcomment{PE}{The plan is to activate winds for frequency bands containing
  strong mesospheric lines, but no tests have yet performed. Describe the wid
  retrieval here, if activated. Otherwise revise text above, we have the always
  same set of variables in \SttVct?}


\section{The a priori vector, \aSttVct{a}}
\label{sec:xa}
%
\lcomment{PE}{Write later. For species we need to decide if we go for the old a
  a proiri database, originally from Bordeaux. Or can we get a better, and much
  newer, alternative? The MIPAS climatology, or from any of all ongoing
  projects?}
\lcomment{PE}{For temperature we need to decide if ECMWF-operational or
  ERA-Interim shall be used as source. And how exactly the transition to MSIS
  is made.}

All sensor correction terms are assumed on average to be zero. Accordingly, the
a priori value of $\Delta\VctStl{o}$, $\Delta\theta$ and $\Delta\nu$ is
throughout set to zero.


\section{A priori uncertainty covariance matrix, \aCvrMtr{a}}
\label{sec:Sa}
%
The setting of \aCvrMtr{a}\ is the most contreversial part of \OEM, see
Sec.~\ref{sec:setup:inverse:xb} for some comments. This subject is not
elaborated further, it is just noticed that for VMR and atmospheric
temperature, in general, there exist no direct measurement on which the
sub-covariance matrices can be based and ad hoc values are applied to a large
extend. In addition, the values selected results in that \aCvrMtr{a}\ in
general should over-estimate variability. Accordingly, \OEM\ is not used in a
strict sense, but the retrievals can still be classified is regularisation
\citep{ungermann2011tomographic} and close to optimal result should be obtained
as long as \aCvrMtr{a}\ is given ``reasonable'' values
\citep{eriksson:analy:00}. On the other hand, the uncertainties of the
instrumental variables are known fairly well and the corresponding parts of
\aCvrMtr{a}\ should be quite accurate.
\\
\lcomment{PE}{Species: Details here depends on retrieval unit selected, and
  text must be written latter. But likely, we use a $\sim$50\% variability,
  with a constrain of a minimum variability defined in \VMR (converted to
  relative values).} \lcomment{PE}{Temperature: For altitudes up to about
  45\,km, a the standard deviation should be $\approx$2\,K. For higher
  altitudes, use fixed values, or does CIRA or MSIS contain tables on natural
  variability that can be used?}

The atmosphere exhibits vertical correlation, see \citet{eriksson:stati:02} for
examples on ozone and temperature correlations. In lack of direct measurements
covering all altitudes, vertical correlations are modelled by parametric
expressions, using a correlation length $l_c$. The correlation is modelled to
follow a Gaussian function,
\begin{equation}
  \label{eq:corr:gau}
  \rho(z_1,z_2) = exp\left(-4[(z_1-z_2)/(l_c(z_1)+l_c(z_2))]^2\right),
\end{equation}
or an exponential one,
\begin{equation}
  \label{eq:corr:exp}
  \rho(z_1,z_2) = exp\left(-2\left|z_1-z_2\right|/(l_c(z_1)+l_c(z_2))\right),
\end{equation}
where $\rho(z_1,z_2)$ is the correlation coefficient between altitudes $z_1$
and $z_2$. Note that the mean of the correlation length at $z_1$ and $z_2$ is
used, that is,  $(l_c(z_1)+l_c(z_2))/2$.\todo{Where do we report $l_c$ used?}

The frequency and pointing off-sets are scalar values, and only a standard
deviation ($\sigma$) for each quantity must be specified. These standard deviations
are set to $\sigma_{\Delta\nu}=?$\,MHz and $\sigma_{\Delta\theta}=?^\circ$, respectively.
The brightness temperature off-sets have been found to be uncorrelated between
spectra, and the variation to have no variation with altitude. This results in
that the covariance matric for $\VctStl{\Delta\VctStl{o}}$ can be written as 
$\sigma_{\VctStl{\Delta\VctStl{o}}}^2\IdnMtr$. The standard deviation of
brightness temperature off-sets is throughout set to ?\,K\todo{Set value for
  all $\sigma$.}.

Any correlations between the different quantities in \SttVct\ are fully
ignored. That is, the total covariance matrix becomes:
\begin{equation}
  \label{eq:sa}
  \aCvrMtr{a} =
  \begin{bmatrix}
    \aCvrMtr{\VctStl{v}_1} & 0 & 0 & 0 & 0 & 0 \\
    0 & \aCvrMtr{\VctStl{v}_2} & 0 & 0 & 0 & 0 \\
    0 & 0 & \aCvrMtr{\VctStl{t}} & 0 & 0 & 0 \\
    0 & 0 & 0 & \sigma_{\VctStl{\Delta\VctStl{o}}}^2\IdnMtr & 0 & 0 \\
    0 & 0 & 0 & 0 & \sigma_{\Delta\theta}^2 & 0 \\
    0 & 0 & 0 & 0 & 0 & \sigma_{\Delta\nu}^2 \\
  \end{bmatrix}.
\end{equation}
The inverse of a block-diagonal matrix of this type is also a block-diagonal
matrix, consisting of the inverse of the individual matrices, and the inverse
of \aCvrMtr{a}\ is calculated in the following manner:
\begin{equation}
  \label{eq:sainv}
  \aCvrMtr{a}^{-1} =
  \begin{bmatrix}
    \aCvrMtr{\VctStl{v}_1}^{-1} & 0 & 0 & 0 & 0 & 0 \\
    0 & \aCvrMtr{\VctStl{v}_2}^{-1} & 0 & 0 & 0 & 0 \\
    0 & 0 & \aCvrMtr{\VctStl{t}}^{-1} & 0 & 0 & 0 \\
    0 & 0 & 0 & \sigma_{\VctStl{\Delta\VctStl{o}}}^{-2}\IdnMtr & 0 & 0 \\
    0 & 0 & 0 & 0 & \sigma_{\Delta\theta}^{-2} & 0 \\
    0 & 0 & 0 & 0 & 0 & \sigma_{\Delta\nu}^{-2} \\
  \end{bmatrix}.
\end{equation}




\section{Forward model parameters, \FrwMdlVct}
\label{sec:b}
%
Please note that static forward model settings, this is, common for all
simulations, are presented already in Sec.~\ref{sec:arts:config}.


\subsection{Pressure grids}
\label{sec:b:pgrid}
%
\dots (also retrieval grids)


\subsection{Spectroscopic data and absorption continua}
\label{sec:b:absdata}
%
\dots


\subsection{Absorption look-up tables}
\label{sec:b:abstable}
%
\dots



\subsection{Radiative transfer}
\label{sec:b:rt}
%
The propagation path used when evaluating Eq.~\ref{eq:rteq} is defined by a set
of points. A point is added at each crossing of the path crosses with a
pressure levels (that is, the pressures defining the vertical grid). Additional
points are inserted to ensure that the distance along the path does not exceed
an user defined threshold. This threshold is denoted $l_\mathrm{max}$ in the
tables of ?\todo{Where?}.
\\
\lcomment{PE}{What about $I_0$?. Is this a static variable? That is, how do we
  handle moon in LOS?}




%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "L2_ATBD"
%%% End: 
