\chapter{L2 Data Product Structure}
\label{chapter:L2_structure}

\section{Accesing Level 2 data}

L2 data can be obtained in two ways:

\begin{enumerate}

\item via a web-api that can be queried and used to get data from
  a specified area and time interval or from specific scans etc.
  Documentation about the web-api is found here: 
  \url{http://odin.rss.chalmers.se/apidocs/index.html} and
  \url{http://odin.rss.chalmers.se/dataaccess}

\item or through downloading monthly L2 product files files over http
  from a file archive (\url{http://odin.rss.chalmers.se/level2_download/}).
  This option can preferably be used if all or a large part of the
  Odin data is of interest.

\end{enumerate}

The web-api provide the L2 data in json format, whereas the
monthly L2 product files have a netCDF format.

\section{Available Level2 products}

A number of \smr\ L2 products are available (\cite{dds}) and
a recommendation about what products to use is available
at \url{http://odin.rss.chalmers.se/dataaccess}.

\section{Monthly L2 product file description}

\subsection{L2 version}
The current L2 data version is v3.0.0.

\subsection{File name convention}

L2 product files are stored in netCDF format
(\url{https://www.unidata.ucar.edu/software/netcdf/}),
and the file name convention of the monthly L2 files is: \\
Odin-SMR\_L2\_\{\textbf{project}\}\_\{\textbf{product}\}\_\{\textbf{year}\}-\{\textbf{month}:02\}.nc, \\
e.g.:\
Odin-SMR\_L2\_ALL-Strat-v3.0.0\_O3-501-GHz-20-to-50-km\_2005-01.nc.\\
where in this case:\\
project = ALL-Strat-v3.0.0 and product = O3-501-GHz-20-to-50-km.
See further in \cite{dds} for available projects and products.


\subsection{File content description}

Each L2 product file contains retrieved VMR or temperature profiles,
where a single profile is associated with a single scan of the atmosphere
by the \smr\ instrument.
The L2 product files covers one month of data, but all retrieved data 
for this month is not necessarily included, as the data is filtered,
and this is further described in ~\ref{sec:data_filtering}.

\subsection*{Variables}

The L2 file contains the variables described below for each retrieved profile: \\

\textbf{Altitude} (array of floats): Altitude of retrieved values (m) \\

\textbf{Apriori} (array of floats): \textit{A priori} profile used in the inversion algorithm ([-] or [K]). \\

\textbf{AVK} (2-D array of floats): Averaging kernel matrix. For gas species, the averaging kernel for relative changes is given ([\%/\%]). For temperature, the unit is [K/K]. \\

\textbf{ErrorNoise} (array of floats): Error due to measurement thermal noise (square root of the diagonal elements of the corresponding error matrix) ([-] or [K]). \\

\textbf{ErrorTotal} (array of floats): Total retrieval error, corresponding to the error due to thermal noise and all interfering smoothing errors (square root of the diagonal elements of the corresponding error matrix) ([-] or [K]). \\

\textbf{GenerationTime} (float): Processing date, given as 'days since 1858-11-17 00:00'. \\ 

\textbf{Lat1D} (float): A scalar representative latitude of the profile [degrees north]. \\ % Say how it has been calculated (mean of 'Latitude'?)

\textbf{Latitude} (array of floats): Approximate latitude of each retrieval value [degrees north]. \\

\textbf{Lon1D} (float): A scalar representative longitude of the profile [degrees east] \\ % Same comment as for 'Lat1D'.

\textbf{Longitude} (array of floats): Approximate longitude of each retrieval value [degrees east]. \\

\textbf{LST} (float): Mean local solar time for the scan [hours]. \\ % Was not included in the previous version. Could make the data more user friendly. 

\textbf{MeasResponse} (array of floats): Measurement response, defined as the row sum of the averaging kernel matrix [-]. \\ % Refer to a document giving recommandations on how it should be used to filter the data.

\textbf{OrbitNum} (int): Odin/SMR orbit number. \\ % Useful? To be discussed.

\textbf{Pressure} (array of floats): Pressure grid of the retrieved profile [Pa]. \\

\textbf{Profile} (array of floats): Retrieved volume mixing ratio [-] or temperature [K] profile  \\ % I think this field should be called 'Profile' instead of 'VMR', in order to be consistent in the case of temperature profile retrieval.

\textbf{ScanID} (int): Unique scan identifier. \\

\textbf{SZA1D} (float): Mean solar zenith angle of the observations used in the retrieval process [degrees]. \\

\textbf{SZA} (array of floats): Approximate solar zenith angle corresponding to each retrieval value [degrees]. \\ % Should be provided? To be discussed. 

\textbf{Temperature} (array of doubles): Estimate of the temperature profile [K] (based on input data from ERA-Interim). \\ % Contrary to what is done in the internal L2 data, I think we should clearly distinguish the retrieved temperature profiles from the ones from the ZPT input. Only the zpt input should be given in this field. When a temperature profile is retrieved, it should be given in the 'Profile' field. Moreover, the source for the ZPT data should be clearly mentioned (either here, or add a reference to a document where the users can find this information).

\textbf{Theta} (array of floats): Estimate of the potential temperature profile [K]. \\

\textbf{Time} (double): Mean time of the scan, given as 'days since 1858-11-17 00:00'. \\ % Not necessary, as the information is already given in the MJD, but I think it can make the data more user friendly to give the time as a string, in addition to the MJD. 

\subsection*{Attributes}

The L2 file contains the following attributes:\\

\textbf{observatiom\_frequency\_mode} (string): Odin/SMR observation frequency mode. \\ % Refer to a table giving all frequency modes with the corresponding frequency ranges and measured products.

\textbf{platform} (string): Instrument name (Odin). \\

\textbf{sensor} (string): Sensor name (SMR). \\

\textbf{inversion\_mode} (string): Inversion mode. \\ % ??

\textbf{level2\_product\_name} (string): L2 product name. \\

\textbf{version\_l1b} (string): Version number of the Level 1b processing chain that was used to create the Level 1b data, in turn used as an input to the Level 2 processor. \\

\textbf{version\_l2} (string): Version number of the Level 2 processing chain that was used to create the Level 2 file. \\

\textbf{time\_coverage\_start} (string):, Mean time of the first scan in the file, given as 'YYYY-MM-DD-hh-mm-ssZ'. \\
                
\textbf{time\_coverage\_end} (string):, Mean time of the last scan in the file, given as 'YYYY-MM-DD-hh-mm-ssZ'. \\


\subsection{File format description}

A description of the L2 netCDF file is given below:

\begin{verbatim}

netcdf Odin-SMR_L2_ALL-Strat-v3.0.0_O3-501-GHz-20-to-50-km_2005-01 {
dimensions:
        time = 5 ;
        level = 28 ;
variables:
        float GenerationTime(time) ;
                GenerationTime:description = "Processing date." ;
                GenerationTime:units = "days since 1858-11-17 00:00" ;
        float Altitude(time, level) ;
                Altitude:description = "Altitude of retrieved values." ;
                Altitude:units = "m" ;
        float Apriori(time, level) ;
                Apriori:description = "A priori profile used in the
                    inversion algorithm." ;
                Apriori:units = "-" ;
        float AVK(time, level, level) ;
                AVK:description = "Averaging kernel matrix." ;
                AVK:units = "%/%" ;
        float ErrorNoise(time, level) ;
                ErrorNoise:description = "Error due to measurement ...";
                ErrorNoise:units = "-" ;
        float ErrorTotal(time, level) ;
                ErrorTotal:description = "Total retrieval error, ...";
                ErrorTotal:units = "-" ;
        float Lat1D(time) ;
                Lat1D:description = "A scalar representative latitude ...";
                Lat1D:units = "degrees north" ;
        float Latitude(time, level) ;
                Latitude:description = "Approximate latitude of each ...";
                Latitude:units = "degrees north" ;
        float Lon1D(time) ;
                Lon1D:description = "A scalar representative longitude ..." ;
                Lon1D:units = "degrees east" ;
        float Longitude(time, level) ;
                Longitude:description = "Approximate longitude of each ..." ;
                Longitude:units = "degrees east" ;
        float LST(time) ;
                LST:description = "Mean local solar time for the scan." ;
                LST:units = "hours" ;
        float MeasResponse(time, level) ;
                MeasResponse:description = "Measurement response, defined ..." ;
                MeasResponse:units = "-" ;
        float Orbit(time) ;
                Orbit:description = "Odin/SMR orbit number." ;
                Orbit:units = "-" ;
        float Pressure(time, level) ;
                Pressure:description = "Pressure grid of the retrieved ..." ;
                Pressure:units = "Pa" ;
        float Profile(time, level) ;
                Profile:description = "Retrieved volume mixing ratio." ;
                Profile:units = "-" ;
        int64 ScanID(time) ;
                ScanID:description = "Satellite time word scan identifier." ;
                ScanID:units = "-" ;
        float SZA1D(time) ;
                SZA1D:description = "Mean solar zenith angle of the ..." ;
                SZA1D:units = "degrees" ;
        float SZA(time, level) ;
                SZA:description = "Approximate solar zenith angle ..." ;
                SZA:units = "degrees" ;
        float Temperature(time, level) ;
                Temperature:description = "Estimate of the temperature ..." ;
                Temperature:units = "K" ;
        float Theta(time, level) ;
                Theta:description = "Estimate of the potential temperature ..." ;
                Theta:units = "K" ;
        float Time(time) ;
                Time:description = "Mean time of the scan." ;
                Time:units = "days since 1858-11-17 00:00" ;

// global attributes:
                :observation_frequency_mode = "1" ;
                :inversion_mode = "stnd" ;
                :level2_product_name = "O3 / 501 GHz / 20 to 50 km" ;
                :date_created = "2020-02-24 T08:31:04Z" ;
                :time_coverage_start = "2005-01-01 T11:12:13Z" ;
                :time_coverage_end = "2005-01-01 T11:30:58Z" ;
                :creator_name = "Donal Murtagh" ;
                :creator_url = "odin.rss.chalmers.se" ;
                :creator_email = "donal.murtagh@chalmers.se" ;
                :address = "412 96 Gothenburg, Sweden" ;
                :institution = "Chalmers University of Technology" ;
                :platform = "Odin" ;
                :sensor = "SMR" ;
                :version_l1b = "8" ;
                :version_l2 = "3.0.0" ;

}
\end{verbatim}


\subsection{Data filtering}
\label{sec:data_filtering}

The montlhy L2 product files are quality filtered, and contain only "valid" data. A retrieved profile
is considered to be valid if 
\begin{itemize}

  \item MinLmFactor \([-]\): The minimum value of the Levenberg - Marquardt factor during the OEM iterations
  is below 2 (for Frequency mode 1 and 2) and 10 (for Frequency mode 8, 13, and 19), and

  \item Residual \([K]\):  The difference between the spectra matching retrieved state and
   used measurement spectra is below 1.5\,K.

Information about the amount of excluded data is found in ~\cite{cadl2data}.
\end{itemize}



\section{Odin web-api}

The L2 part of the Odin web-api contains a number of endpoints
(\url{http://odin.rss.chalmers.se/apidocs/index.html#/level2}),
that provides three different L2 data objects denoted as
L2, L2i, and L2anc. The L2 object contains the main result of
the retrieval calculation (e.g. retrieved profile and averaging kernels), 
the L2i object contains obtained offsets and residual etc. from the
retrieval calculation, and the L2anc object contains ancillary data.
The endpoints provide the L2 data in JSON format
and the content of these three objects are described below.

\subsection{L2}
\input{../IODD/l2_format}
\subsection{L2i}
\input{../IODD/l2i_format}
\subsection{L2anc}

\begin{itemize}

    \item FreqMode (\emph{int}): \smr\ frequency mode of the observation
    \item InvMode (\emph{string}): Inversion mode
    \item LST (\emph{double}): Mean local solar time for the scan [hours].
    \item Lat1D (\emph{double}): A scalar representative latitude of
      the retrieval
    \item Latitude (\emph{array of doubles}): Approximate latitude of each
      retrieved value
    \item Lon1D (\emph{double}): A scalar representative longitude of
      the retrieval
    \item Longitude (\emph{array of doubles}): Approximate longitude of each
      retrieved value
    \item MJD (\emph{double}): Mean modified julian date of the observations
      used in the retrievel
    \item Orbit (\emph{double}): Orbit number
    \item Pressure (\emph{array of doubles}): Pressure grid of the retrieved
      species [Pa]
    \item SZA  (\emph{double}): Mean solar zenith angle of the observations
      used in the retrieval process [degrees]
    \item SZA1D (\emph{double}): Mean solar zenith angle of the observations
      used in the retrieval process [degrees]
    \item ScanID (\emph{int}): Satellite time word scan identifier
    \item Theta (\emph{array of doubles}): Estimate of the potential temperature
      profile [K]

\end{itemize}
