\chapter{In Progress}
\label{chapter:in_progress}



\section{Work Packages}

\begin{itemize}


\item{\bf WP 2.2.1.2 - Implementation}

The implementation of the bulk reprocessing system has begun. This WP is heavily dependent on the analysis of the prototype system and the DDS produced by it.

\item{\bf WP 2.2.1.3 - Testing}

Testing of the bulk reprocessing system will be performed continuously in parallel with the implementation. 


\item{\bf WP 2.2.3 - Construction of prototype system for bulk reprocessing}

A prototype exists and is running at this moment on the VDS to produce the DDS. It is expected that this WP will remain open until the construction of the final processing system is well under way.  


\item{\bf WP 2.2.4 - Generate DDS}

The production of the DDS from the VDS in ongoing. Currently results have been produced for frequency modes 1 and 2, with 13 and 21 next in line. The generation of the DDS is motivated for two fundamentaly different reasons: for analysing the data and optimizing the algorithms and calibrations; and for testing and improving the prototype for the bulk reprocessing system. The first task will take longer to finish than the second.


\item{\bf WP 3.2 - Design of database model}

The L1 database is already in place and filled with appropriate data. It is a postgreSQL database system. The L2 database is being implemented in mongoDB, and is expected to show improved performance with respect to speed in comparison to the postgreSQL system.  


\item{\bf WP 5.1 - Project management plan (PMP) as activities/tasks/personnel change}

A second version of of the PMP has been delivered, reflecting schedule changes, but
this work package remains active until the project is finished, as the PMP will be
updated upon any notable changes in the project management.



\item{\bf WP 5.2.4 - Tri-Monthly Management Report 4}

This document is the fourth tri-monthly management report. 


\end{itemize}


\section{Deliverables}


\begin{itemize}


\item{\bf D-7 - Diagnostics dataset (DDS)}

The production of the DDS from the VDS is ongoing and running on the prototype reprocessing system. See WP 2.2.4.


\item{\bf D-8 - Product Validation and Evolution Report (PVR v.1)}

The DDS will be evaluated and the results will be reported in them PVR document.


\end{itemize}
