\chapter{Progress since last report}
\label{chapter:progress}


\section{Work Packages}

\begin{itemize}

\item{\bf WP 1.1.2.3 - Comparison of calibration versions}

This WP deals with comparison of calibration version 7 (v.7) and 8 (v.8) data. v.8
builds upon v.7 calibration routine, but is extended and more advanced. v.8 is scan-
based whereas v.7 is orbit based, and v.8 aims to remove ripples seen in the spectra
in the upper part of the scan in the v.7 data.
Production of comparison plots of data from the two versions has been made. On many
occasions the data from the two versions differs only marginally, but v.8 data clearly
shows better performance for measurements in the upper part of scan.


\item{\bf WP 1.1.3 - Technical note (L1 REV TN)}

Due to the large amount of material, what was planned as one technical note in WP
1.1.3 has been split into two documents: the L1 ATBD, and a document called "Odin
SMR Calibration Accuracy". Both of these documents have been delivered to ESA. 


\item{\bf WP 1.2.1.1 - Coverage in Time and Observation mode}

Detailed interactive statistics can now be found on the Odin web page (odin.rss.chalmers.se). 


\item{\bf WP 1.2.2 - Documentation in L1 Completeness Analysis Document (L1 CAD)
}

It was deemed that the webpage was a better form for presenting the data regarding the completeness of the data than a static document. In view of the dynamic nature of the data, and the constant desire for improvement in the form of presentation. The webpage is up, and most functions are satisfactory, some details remain to be fixed e.g. only successfull scans are presented, without any information regarding what the planned scans were. This makes it hard to judge the performance of the instrument. This will be taken care of during the project life, but right now it is not prioritized. 


\item{\bf WP 1.3.1 - Selection of Representative Orbits (Good/Bad)}

The dataset comprising the VDS has been chosen with regard to several factors, e.g. longitude-latitude coverage, and the availability of colocation, but also with regard taken to the data being of "good/bad" quality. Details regardig this can be found in the VDS technical note. 


\item{\bf WP 1.3.2 - Packaging of data for testing}

The subset of data which makes up the VDS has been copied from the complete database and put into a database exclusively for the VDS, and for the subsequent generation of the DDS. 


\item{\bf WP 1.3.3 - Inventory of other datasets}

Careful colocations have been made with a number of other datasources, e.g. ballon soundings, in situ data, and other satellite instruments. The details regarding this can be found in the VDS technical note.


\item{\bf WP 1.3.4 - Level1 verification dataset technical note}

The VDS has been documented and described in a document called "Verification Dataset: Technical Note" that has been delivered to ESA.


\item{\bf WP 2.1.3 - Write Requirements Baseline Document (RB)}

The RB document has been delivered to ESA. It contains all relevant information regarding what the finished L2 product will be, what is needed to produce it in terms of algorithms and data, and what is needed to verify the product. 


\item{\bf WP 5.2.2 - Tri-Monthly Management Report 2}

This document is the second tri-monthly management report.


\end{itemize}



\section{Deliverables}

\begin{itemize}

\item{\bf D-1 - Project Management Plan (PMP)}

An updated version of the PMP is delivered along with this document (TMMR2).
The updated version is changed with respect to the delivery date of D-7, the diagnostics dataset (DDS). Due to a larger DDS than originally planned for the projected computing time had to be extended, and the new date for delivery of the DDS will be 2016-09-30.  

\item{\bf D-2 - Tri-Monthly Management Report 2 (TMMR2)}

This document is the second tri-monthly management report.

\item{\bf D-3 - Requirements Baseline (RB) Document}

The RB document has been delivered to ESA. It contains all relevant information regarding what the finished L2 product will be, what is needed to produce it in terms of algorithms and data, and what is needed to verify the product. 


\item{\bf D-4 - L1 Completeness Analysis Document (L1-CAD)}

It was deemed that the webpage was a better form for presenting the data regarding the completeness of the data than a static document. In view of the dynamic nature of the data, and the constant desire for improvement in the form of presentation. The webpage is up, and most functions are satisfactory, some details remain to be fixed e.g. only successfull scans are presented, without any information regarding what the planned scans were. This makes it hard to judge the performance of the instrument. This will be taken care of during the project life, but right now it is not prioritized. 


\item{\bf D-5 - Verification Dataset (VDS)}

A representative and colocated VDS has been selected from the entire dataset. It amounts to roughly 3 \% of the complete dataset. The VDS has been documented and described in a document called "Verification Dataset: Technical Note" that has been delivered to ESA. What remains to be done regardig the VDS is to get a hold of all the external colocated data.


\item{\bf D-16 - L1 Review Technical Note}

Due to the large amount of material, what was planned as one technical note in WP
1.1.3 has been split into two documents: the L1 ATBD, and a document called "Odin
SMR Calibration Accuracy". Both of these documents have been delivered to ESA. 

\end{itemize}

