\Chapter{Data formats for the instruments}
\label{sec:dataformat}
\subsection{Aura/MLS}
The MLS data collocated with Odin/SMR is accessible through the Odin REST
API, see section~\ref{sec:api}. The data is returned as a JSON object with the
following attributes:
\begin{itemize}
    \item data\_fields \emph{(Object)}: Object containing the data under the
        following keys:
        \begin{itemize}
            \item AscDescMode \emph{(Int)}: 0 for ascending, 1 for descending
                measurement mode
            \item Convergence \emph{(Double)}: Convergence of the ...
                \todo{what?}
            \item <Species> \emph{(Array of doubles)}: Concentration profile
                for <Species>
            \item <Species>Precision \emph{(Array of doubles)}: Precision of
                the concentration profile for <Species>
            \item L2gpValue \emph{(Array of doubles)}: Level 2 ... \todo{what?}
            \item L2gpPrecision \emph{(Array of doubles)}: Precision of the ...
                Level 2 ... \todo{what?}
            \item Quality \emph{(Double)}: Quality of the ... \todo{what?}
            \item Status \emph{(Int)}: Status code ...
        \end{itemize}
    \item geolocation\_fields \emph{(Object)}: Object containing the
        geolocation of the data under the following keys:
        \begin{itemize}
            \item ChunkNumber \emph{(Int)}: Number of chunks...
            \item Latitude \emph{(Double)}: Latitude for the measurement
            \item Longitude \emph{(Double)}: Longitude for the measurement
            \item LineOfSightAngle \emph{(Double)}: Line of sight angle for the
                measurement
            \item LocalSolarTime \emph{(Double)}: Local solar time for the
                measurement
            \item MJD \emph{(Double)}: Time for the measurement in MJD
            \item Time \emph{(Double)}: Time for the measurement in ...
                \todo{what?}
            \item OrbitGeodeticAngle \emph{(Double)}: The geodetic angle of the
                orbit at the time of measurement
            \item SolarZenithAngle \emph{(Double)}: The solar zenith angle for
                the measurement
            \item Pressure \emph({Array of doubles}): Pressure profile for the
                measurement
        \end{itemize}
\end{itemize}

\subsection{ENVISAT/MIPAS}
The MIPAS data collocated with Odin/SMR is accessible through the Odin REST
API, see section~\ref{sec:api}. The data is returned as a JSON object with the
following attributes:
\begin{itemize}
    \item MJD \emph{(Double)}: Time in MJD for the measurement
    \item altitude \emph{(Array of doubles)}: Altitudes for the measurement
    \item laitutde \emph{(Double)}: Latitude for the measurement
    \item longitude \emph{(Double)}: Longitude for the measurement
    \item los \emph{(Array of doubles)}: Line of sight angles for the
        measurement
    \item sza \emph{(Double)}: Sun zenith angle for the measurement
    \item time \emph{(Double)}: Time of the measurement in ... \todo{what?}
    \item pressure \emph{(Array of doubles)}: Pressure profile for the
        measurement
    \item temperature \emph{(Array of doubles)}: Temperature profile for the
        measurement
    \item akm\_diagonal \emph{(Array of doubles)}: Diagonal something ...
        \todo{what?}
    \item chi2 \emph{(Double)}: $\chi^2$ of ... \todo{what?}
    \item dof \emph{(Double)}: Degrees of freedom ... \todo{or? why double?}
    \item eta \emph{(Array of doubles)}: Diagonal something with $\eta$...
        \todo{what?}
    \item eta\_indices \emph{(Array of ints)}: Indices of something ...
        \todo{what?} with $\eta$
    \item rms \emph{(Double)}: Root mean square error of something ...
        \todo{what?}
    \item geo\_id \emph{(Array of strings)}: Some sort of id ...
        \todo{what?} with $\eta$
    \item sub\_id \emph{(Array of strings)}: Some other sort of id ...
        \todo{what?} with $\eta$
    \item target \emph{(Array of doubles)}: Target of something ...
        \todo{what?}
    \item target\_noiser\_error \emph{(Array of doubles)}: Noise error of
        something target something ... \todo{what?}
    \item visibility \emph{(Array of ints)}: Visibility of something ...
        \todo{what?}; 0 indicates obusfacted, 1 indicates visible
    \item vr\_akdiag \emph{(Array of doubles)}: Something diagonals ...
        \todo{what?}
    \item vr\_col \emph{(Array of doubles)}: Something columns ... \todo{what?}
    \item vr\_row \emph{(Array of doubles)}: Something rows ... \todo{what?}
\end{itemize}

\subsection{ISS/JEM/SMILES}
The SMILES data collocated with Odin/SMR is accessible through the Odin REST
API, see section~\ref{sec:api}. The data is returned as a JSON object with the
following attributes:
\begin{itemize}
    \item data\_fields \emph{(Object)}: Object containing the data under the
        following keys:
        \begin{itemize}
            \item AOSUnitNum \emph{(Double)}: Number of the ...
                \todo{what?}
            \item Apriori \emph{(Array of doubles)}: Apriori profile of ...
                \todo{what?}
            \item AprioriError \emph{(Array of doubles)}: Apriori error of ...
                \todo{what?}
            \item AveragingKernel \emph{(Two-dimensional array of doubles)}:
                Averaging kernel used in the ... \todo{what?}
            \item Baseline<N> \emph{(Array of doubles)}: Baselines of ...
                \todo{what?}; N=0--3
            \item Baseline<n>Precision \emph{(Array of doubles)}: Precision of
                baseline of ... \todo{what?}; N=0--3
            \item Convergence \emph{(Double)}: Measurement of the convergence
                of the ... \todo{what?}
            \item CorrLength \emph{(Double)}: Correlation length of the ...
                \todo{what?}
            \item CostFunctionY \emph{(Array of doubles)}: Cost function ...
                \todo{what?}
            \item CostFunctionYAll \emph{(Double)}: Cost function ...
                \todo{what?}
            \item DifferenceY \emph{(Array of doubles)}: Differences ...
                \todo{what?}
            \item DifferenceYAll \emph{(Double)}: Differences ...
                \todo{what?}
            \item FOVInterference \emph{(Int)}: Field-of-view interference flag
                \todo{or?}
            \item InformationValue \emph{(Array of doubles)}: Value of ...
                \todo{what?}
            \item L2Value \emph{(Array of doubles)}: Level 2 ... \todo{what?}
            \item L2Precision \emph{(Array of doubles)}: Precision of the ...
                Level 2 ... \todo{what?}
            \item MaxNumIteration \emph{(Int)}: Maximum number of iterations
                for the ... \todo{what?} algorithm
            \item NumIterationPerform \emph{(Int)}: Number of iteration in the
                ... \todo{what?} algorithm
            \item MeasurementError \emph{(Array of doubles)}: Estimate of the
                measurement error of the ... \todo{what?}
            \item PrecisionWOSignal \emph{(Array of doubles)}: Precision of
                the ... \todo{what?}
            \item Pressure \emph({Array of doubles}): Pressure profile for the
                measurement
            \item Temperature \emph({Array of doubles}): Temperature profile
                for the measurement
            \item RadianceResidualMax \emph{(Double)}: Maximum of the residual
                radiance of the ... \todo{what?}
            \item RadianceResidualMean \emph{(Double)}: Mean of the residual
                radiance of the ... \todo{what?}
            \item RadianceResidualRMS \emph{(Double)}: Root mean square error
                of the residualvradiance of the ... \todo{what?}
            \item SeqCount \emph{(Int)}: Number of ... \todo{what?}
            \item SmoothingError \emph({Array of doubles}): Error from the
                smoothing of ... \todo{what?}
            \item VerticalResolution \emph({Array of doubles}): Vertical
                resolution of the profiles
            \item WaterVapor \emph({Array of doubles}): Profile of the water
                vapour concentration
        \end{itemize}
    \item geolocation\_fields \emph{(Object)}: Object containing the
        geolocation of the data under the following keys:
        \begin{itemize}
            \item AscendingDescending \emph{(Int)}: 0 for ascending, 1 for
                descending measurement mode
            \item Latitude \emph{(Double)}: Latitude for the measurement
            \item Longitude \emph{(Double)}: Longitude for the measurement
            \item LineOfSightAngle \emph{(Double)}: Line of sight angle for the
                measurement
            \item MJD \emph{(Double)}: Time for the measurement in MJD
            \item LocalTime \emph{(Double)}: Local time of day for the
                measurement in ... \todo{what?}
            \item LocalTime \emph{(Double)}: Time for the measurement in ...
                \todo{what?}
            \item TimeUTC \emph{(String)}: Date and time for the measurement in
                UTC as a string in ISO format
                \todo{what?}
            \item Reserved \emph{(Int)}: Reserved ... \todo{what?}
            \item SolarZenithAngle \emph{(Double)}: The solar zenith angle for
                the measurement
            \item Altitude \emph({Array of doubles}): Altitudes (km) for the
                measured profiles
        \end{itemize}
\end{itemize}


\subsection{Odin/OSIRIS}
...


\subsection{Meteor3M/SAGE III}
The SAGE III data collocated with Odin/SMR is accessible through the Odin REST
API, see section~\ref{sec:api}. The data is returned as a JSON object with the
following attributes:
\begin{itemize}
    \item FileName    \emph{(String)}: Filename is the same as in the original
        data set
    \item Instrument  \emph{(String)}: Name of the instrument
    \item EventType   \emph{(String)}: Solar or Lunar (only Solar included in
        VDS)
    \item MJDStart    \emph{(Double)}: Start time in MJD for the measurement
    \item MJDEnd      \emph{(Double)}: End time in MJD for the measurement
    \item LatStart    \emph{(Double)}: Start latitude for the measurement
    \item LatEnd      \emph{(Double)}: End latitude for the measurement
    \item LongStart   \emph{(Double)}: Start longitude for the measurement
    \item LongEnd     \emph{(Double)}: End longitude for the measurement
    \item Pressure    \emph{(Array of doubles)}: Pressure profile for the
        measurement
    \item Temperature \emph{(Array of doubles)}: Temperature profile for the
        measurement
    \item <Species>   \emph{(Array of triplets of doubles)}: Profile for
        <Species> for the measurement; each row is a triplet containing
        concentration, uncertainty, and a quality bit
        %flag\footnote{see \cite{SAGEIII_DPUG} for details!} \\
\end{itemize}
\TODO{Add line of sight to SAGE III api?}
