\chapter{Conclusions}
\label{chapter:conclusions}
\section{General comments}


Table~\ref{table:level2:v3} shows an overview of the characteristics of the
main data products from the updated \smr\ processing chain for the frequency
modes investigated in Chapter~\ref{chapter:comparisons}. The vertical coverage
has been defined as the altitude interval where the \emph{measurement response}
for the retrievals is $>0.8$. The vertical resolution and precision are
calculated from the $\pm 2\sigma$ percentiles of the \emph{full widths at half
maximum} of the averaging kernels and the \emph{total error} respectively.

The results presented in Table~\ref{table:level2:v3} should be compared with
the results for the old processing chain, presented in
Tables~\ref{table:level2} and~\ref{table:level2b}.

\begin{table}[tbhp]
    \centering
    \caption{Characteristics of \smr\ Level2 main data products.}
    \label{table:level2:v3}
    \scalebox{0.965}{%
    \begin{tabular}{lrrrr}
        \toprule
        \textbf{Product} & \textbf{Frequency} & \textbf{Vertical}   & \textbf{Vertical}     & \textbf{Precision} \\
                         & \textbf{{[}GHz{]}} & \textbf{coverage}   & \textbf{resolution}   &                    \\
        \midrule  % FM 01:
        \chem{O_{3}}     & 501.5 (FM~01)      & 19--53\,km          & 5.3--5.5\,km          & 0.70--0.85\,ppmv \\
        \chem{ClO}       & 501.3 (FM~01)      & 18--58\,km          & 5.3--5.6\,km          & 0.07--0.08\,ppbv \\
        \chem{N_{2}O}    & 502.3 (FM~01)      & 15--62\,km          & 4.4--5.6\,km          & 5.00--6.10\,ppbv \\

        \midrule  % FM 02:
        \chem{O_{3}}     & 544.9 (FM~02)      & 17--77\,km          & 2.9--4.9\,km          & 0.18--0.26\,ppmv \\
        \chem{HNO_{3}}   & 544.4 (FM~02)      & 20--61\,km          & 5.0--7.2\,km          & 0.24--0.28\,ppbv \\
        Temperature      & 544.9 (FM~02)      & 21--64\,km          & 7.5--7.9\,km          & 1.13--2.18\,K \\

        \midrule  % FM 08:
        \chem{O_{3}}     & 489.2 (FM~08)      & 16--71\,km          & 4.1--6.0\,km          & 0.29--0.39\,ppmv \\
        \chem{H_{2}O}    & 489.5 (FM~08)      & 19--78\,km          & 4.0--4.7\,km          & 0.32--0.55\,ppmv \\

        \midrule  % FM 13:
        \chem{O_{3}}     & 556.9 (FM~13)      & 44--80\,km          & 6.1--6.7\,km          & 0.19--0.24\,ppmv \\
        \chem{H_{2}O}    & 556.7 (FM~13)      & 44--110\,km         & 4.3--4.8\,km          & 0.06--0.33\,ppbv \\
        Temperature      & 556.9 (FM~13)      & 44--95\,km          & 5.3--5.6\,km          & 1.85--2.73\,K \\

        \midrule  % FM 19:
        \chem{O_{3}}     & 556.9 (FM~19)      & 43--81\,km          & 5.8--6.6\,km          & 0.19--0.24\,ppmv \\
        \chem{H_{2}O}    & 556.7 (FM~19)      & 43--109\,km         & 4.1--4.4\,km          & 0.13--0.26\,ppbv \\
        Temperature      & 556.9 (FM~19)      & 43--93\,km          & 5.2--5.5\,km          & 1.72--2.65\,K \\
        \bottomrule
    \end{tabular}}
\end{table}

\section{Recomendations}
\subsection{Ozone}
The best ozone product for general use is that from FM~02.  The new product is
in much better agreement with other instruments and does not seem to show any
temporal drifts although a thorough analysis of drifts has not yet been
attempted.  The other ozone products show some biases but could be used in
conjunction with simultaneously measured products if account is taken of this.

\subsection{Chlorine Monoxide}
The \chem{ClO} product should be reliable until 2009 and after January 2018.
Unfortunately a gradually deteriorating instrument malfunction has disturbed
the intervening period.  We will attempt to correct for this as soon as
possible

\subsection{Water vapour}
Water vapour continues to be a difficult product. the FM~08 product has
improved and is stable over time although shows a 15~\% low bias compared to
MLS.  For the upper stratosphere and mesosphere FM~13 appears to be the best
product, but also shows a small bias of around 15~\% and som variability above
70~km that needs further investigation.

\subsection{Temperature}
FM~02 provides stratospheric temperatures with a possible increasing bias with
altitude. There does not appear to be any systematic drift with time. For
higher altitudes FM~13 can be used with knowledge of a possible cold bias of
3--5~K
