\chapter{In Progress}
\label{chapter:in_progress}



\section{Work Packages}

\begin{itemize}

\item{\bf WP 1.1.1.3 - Comparison with simulated data}

Work on this WP is still ongoing, details can be found in TMMR1.


\item{\bf WP 1.2.1.2 - Categorisation of Data (Good/Bad)}

The aim of this WP is to Categorise data into different quality levels, identify how quality levels for a scan be defined, and produce a table that describes how large fraction of the scans that falls into the different quality categories for each month of the complete mission. A rudimentary system for putting quality flags on spectra and scans is in place, but the final quality level system should not be finished before the VDS has been processed into the DDS, and an evaluation of the quality has been done.   


\item{\bf WP 2.2.1.1 - Design}

WP's 2.2.1.1, 2.2.1.2, and 2.2.1.3 are run in parallel in an agile process. The work on design and implementation is nearing completion. 

\item{\bf WP 2.2.1.2 - Implementation}

See WP 2.2.1.1.

\item{\bf WP 2.2.1.3 - Testing}

See WP 2.2.1.1.


\item{\bf WP 5.1 - Project management plan (PMP) as activities/tasks/personnel change}

A second version of of the PMP has been delivered, reflecting schedule changes, but this work package remains active until the project is finished, as the PMP will be updated upon any notable
changes in the project management.

\item{\bf WP 5.2.3 - Tri-Monthly Management Report 3}

This document is the third tri-monthly management report. 


\end{itemize}


\section{Deliverables}

Currently there are no deliverables in progress, and none are planned until autumn.