\chapter{Comparisons with other instruments}
\label{chapter:comparisons}
This chapter contains comparisons between the Level2 data products of \smr\
and collocated measurements from various instruments as accumulated in the
Validation Data Set (VDS)~(\cite{VDS:2016}). The initial VDS covered the
period 2003--2015 but this has been extended to 2019. This implies that the
DDS has also been extended to the same period. The comparisons are organised
primarily on frequency mode, secondarily on data product. For each product,
average deviations from the various instruments at different altitudes are
investigated, followed by an investigation of the overall correlation between
the measurements and both \smr~v2 and~v3. The \smr\ measurement response and
averaging kernels are also investigated for each data product~\cite{atbdl2} .
This has been carried out for the main data products i.e. those from FM~01,
FM~02, FM~08, FM~13 and FM~19. The other science mode products require
separate treatment since they must be compared to special products from the
other instruments. They have therefore been excluded from the project and
this report . However many of the ``lessons learnt'' in the project will be
useful. From the correlation some key performance indicators are obtained and
presented in a table. Of these the most important are the \emph{Pearson
correlation} and \emph{slope}, both of which should be as close as possible
to unity for a good match, and the \emph{intercept}, which should ideally be
zero.
